---
title: "EconHw3"
author: "Trevor Freeland"
date: "April 23, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, comment = NA, warning = F)
```

```{r}
library(tidyverse)
library(readxl)
library(pander)
```

```{r}
table6 <- read_xls("~/R/Econometrics/Tables/Table7_6.xls")
table6 <- table6[-(1:10),]
colnames(table6) <- table6[1,]
table6 <- table6[-1,]
table6[] <- sapply(table6[], as.numeric)
```


```{r}
day8 <- read.csv("~/R/Econometrics/Tables/Day08.csv")
```

Tuesday Stuff:

```{r}
ggplot(day8, aes(x = log(X3), y = log(Y))) + geom_point() + stat_smooth(method = "lm")
day8.lm <- lm(log(Y)~log(X2) + log(X3), data = day8)
day8.lm2 <- lm(log(Y)~log(X2) + log(X3), data = day8)
AIC(day8.lm2)
AIC(day8.lm)
```

##1

A: False

B: False

C: False

D: True

##2

###(A)

$\alpha_2$ = -2228 -> Holding all else constant, if the average wholesale price of roses went up by $1/dozen, on average the number of roses sold would decrease by 2227.

$\alpha_3$ = 1251 -> Holding all else constant, if the average wholesale price of carnations went up by $1/dozen, on average the number of roses sold would increase by 1251.

$\alpha_4$ = 6.3 -> Holding all else constant, if the average weekly family disposable income grew by $1/week, on average the number of roses sold would increase by 6.

$\alpha_5$ = -197 -> Holding all else constant, on average the quantity of roses sold in a quarter decreased by 197 roses per quarter starting in 1971-III. 

See table below for full model summary.

```{r}
rose.lm <- lm(Y~X2 + X3 + X4 + X5, data = table6)
pander(summary(rose.lm))
```

###(B)

$\beta_2$ = -1.17 -> Holding all else constant, if the average wholesale price of roses went up by 1% per dozen, on average the number of roses sold would decrease by 1.17%.

$\beta_3$ = .74 -> Holding all else constant, if the average wholesale price of carnations went up by 1% per dozen, on average the number of roses sold would increase by .74%.

$\beta_4$ = 1.15 -> Holding all else constant, if the average weekly family disposable income grew by 1% per week, on average the number of roses sold would increase by 1.15%.

$\beta_5$ = -.18 -> Holding all else constant, on average the quantity of roses sold in a quarter decreased by 3% roses per quarter starting in 1971-III. 

See table below for full model summary.

```{r}
rose.loglm <- lm(log(Y)~log(X2) + log(X3) + log(X4) + X5, data = table6)
summary(rose.loglm)
```

###(C)

Own-Price: We would expect as the price of the product increased, demand would decrease. (Holding everything else constant)

Cross-Price: We would expect as the price of a competitor's product increases, the demand for our product would increase. (Holding everything else constant)

Income Elasticity: We would expect as our buyers had more money to spend, that the demand for our product would increase. (Holding everything else constant)

Our results concur with our expectations. 

###(D)



###(E)



###(F)

##3

##4

##5

##Challenge Problem