---
title: "EconHw3"
author: "Trevor Freeland"
date: "April 23, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, comment = NA, warning = F)
```

```{r}
library(tidyverse)
library(readxl)
library(pander)
```

```{r}
table6 <- read_xls("~/R/Econometrics/Tables/Table7_6.xls")
table6 <- table6[-(1:10),]
colnames(table6) <- table6[1,]
table6 <- table6[-1,]
table6[] <- sapply(table6[], as.numeric)
```

##1

A: False

B: False

C: False

D: True

##2

###(A)

$\alpha_2$ = -2228 -> Holding all else constant, if the average wholesale price of roses went up by $1/dozen, on average the number of roses sold would decrease by 2227.

$\alpha_3$ = 1251 -> Holding all else constant, if the average wholesale price of carnations went up by $1/dozen, on average the number of roses sold would increase by 1251.

$\alpha_4$ = 6.3 -> Holding all else constant, if the average weekly family disposable income grew by $1/week, on average the number of roses sold would increase by 6.

$\alpha_5$ = -197 -> Holding all else constant, on average the quantity of roses sold in a quarter decreased by 197 roses per quarter starting in 1971-III. 

See table below for full model summary.

```{r}
rose.lm <- lm(Y~X2 + X3 + X4 + X5, data = table6)
pander(summary(rose.lm))
```

###(B)

$\beta_2$ = -1.17 -> Holding all else constant, if the average wholesale price of roses went up by 1% per dozen, on average the number of roses sold would decrease by 1.17%.

$\beta_3$ = .74 -> Holding all else constant, if the average wholesale price of carnations went up by 1% per dozen, on average the number of roses sold would increase by .74%.

$\beta_4$ = 1.15 -> Holding all else constant, if the average weekly family disposable income grew by 1% per week, on average the number of roses sold would increase by 1.15%.

$\beta_5$ = -.18 -> Holding all else constant, on average the quantity of roses sold in a quarter decreased by 3% roses per quarter starting in 1971-III. 

See table below for full model summary.

```{r}
rose.loglm <- lm(log(Y)~log(X2) + log(X3) + log(X4) + X5, data = table6)
summary(rose.loglm)
```

###(C)

Own-Price: We would expect as the price of the product increased, demand would decrease. (Holding everything else constant)

Cross-Price: We would expect as the price of a competitor's product increases, the demand for our product would increase. (Holding everything else constant)

Income Elasticity: We would expect as our buyers had more money to spend, that the demand for our product would increase. (Holding everything else constant)

Our estimates in our model concur with our expectations. 

###(D)

Own-Price:

Cross-Price:

Income Elasticity:

###(E)

I would choose to use the log-log model for our analysis. Especially in this setting, it seems we are very interested in Own-Price/Cross-Price/Income-Elasticity, which are all far easier to find, understand and interpret using a log-log model. Also knowing the scales of both of the variables are extremely large, typically you would use a log-log model instead of a linear model. 

###(F)

We want to test to see if the estimate for log(X2) = 1. So lets make a test stat:

$$t = \frac{-1.17 - 1}{.49} = -4.43$$

Even before calculating the actual p-value we know that with a t stat over 4 standard deviations away from the hypothesized value that it is very unlikely to happen. THe p-value we get when testing out t-stat we get a p-value smaller than .001, which is strong evidence that our null hypothesis is wrong, and so we do not believe that demand is own-price unit elastic.

```{r}
t.stat <- (-1.17 - 1)/.49
p <- pt(t.stat, 11)
```

##3

##4

##5

##Challenge Problem